---
title: "Final Plots for Dengue"
author: "Toby Koch"
date: "2023-07-11"
output: pdf
---

LOAD IN DATA
```{r setup, include=FALSE}
rm(list = ls())

source("scripts/header_code.R")
source("scripts/functions.R")
load("data_input/df_yr.RData")
load("data_input/case_paho_yr.RData")

```

SOME ADDITIONAL DATA CLEANING

```{r}
df_yr_train = df_yr %>%
  filter(Country %in% c("Puerto Rico", "Dominican Republic", "Jamaica")) %>%
  mutate(country_mod = Country)

df_yr_test = df_yr %>%
  mutate(Cases = ifelse(Country %in% c("Cuba", "Haiti"),
                         NA,
                       Cases))

```

TRAVEL INCIDENCE SENSITIVITY ANALYSIS
```{r}
data <- df_yr_train  %>%
  select(Cases, travel_incidence_dot, gdp_cat, year0, Pop)
   

# List of transformations to test
trans <- seq(0.001,0.02, 0.001)

#calculate the transformatins
for (i in trans) {
  col_name <- paste0("log_incidence_", i)

  # Mutate new column
  data <- data %>% 
    mutate(!!col_name := log(travel_incidence_dot + i))
}


# Function to fit NB regression and get AIC score
fit_and_get_aic <- function(data, formula) {
  model <- glm.nb(formula, data = data)
  return(AIC(model))
}

# List to store AIC scores
aic_scores <- c()

# Loop through each log-transformed variable and fit Poisson regression model
for (i in seq_along(trans)) {
  col_name <- paste0("log_incidence_", trans[i])
  formula <- as.formula(paste("Cases ~", col_name, "+ gdp_cat + year0 + offset(log(Pop))"))
  aic_score <- fit_and_get_aic(data, formula)
  aic_scores[i] <- aic_score
}

aic_trans = data.frame(trans,aic_scores)

plot(trans, aic_scores)


data <- df_yr_train  %>%
  select(Cases, travel_incidence_dot, gdp_cat, year0, Pop)
   

# List of transformations to test
trans <- seq(0.001,0.03, 0.001)

#calculate the transformatins
for (i in trans) {
  col_name <- paste0("log_incidence_", i)

  # Mutate new column
  data <- data %>% 
    mutate(!!col_name := log(travel_incidence_dot + i))
}


# Function to fit NB regression and get AIC score
fit_and_get_aic <- function(data, formula) {
  model <- glm.nb(formula, data = data)
  return(AIC(model))
}

# List to store AIC scores
aic_scores <- c()

# Loop through each log-transformed variable and fit Poisson regression model
for (i in seq_along(trans)) {
  col_name <- paste0("log_incidence_", trans[i])
  formula <- as.formula(paste("Cases ~", col_name, "+ gdp_cat + year0 + offset(log(Pop))"))
  aic_score <- fit_and_get_aic(data, formula)
  aic_scores[i] <- aic_score
}

aic_trans = data.frame(trans,aic_scores)

plot(trans, aic_scores)



```


```{r}
modnb_yr_log_gdp2.2 = glm.nb(Cases ~ travel_incidence_dot_log + gdp_cat + year0 + offset(log(Pop)) , 
               data = df_yr_train)

mod_final = modnb_yr_log_gdp2.2


#create df of estimates
df_mod_final = tidy(mod_final, 
                    conf.int = T, 
                    exponentiate = T) %>%
  mutate_if(is.numeric, ~round(.,2))

#change the model output data frame text to be more readable for the final version
df_mod_final = 
  df_mod_final %>% 
  rename("IRR" = "estimate") %>%
  filter(term != "(Intercept)") %>%
  mutate(term = case_when(term == "travel_incidence_dot_log" ~ "log(Travel Incidence)",
                          term == "year0" ~ "Year",
                          term == "gdp_catupper" ~ "GDP:Upper Income")
          )

write.csv(df_mod_final, "data_output/df_mod_estimates.csv")
```


CHANGE CASES FOR CUBA TO SYNDROMIC TO ACCOUNT FOR ENHANCED SURVEILLANCE EFFORTS
```{r}
df_yr = df_yr %>%
  mutate(`Travel Cases` = if_else(!is.na(syndromic_cases),
                                   syndromic_cases,
                                   `Travel Cases`)) %>%
  mutate(travel_incidence_dot = round(`Travel Cases`/pass_dot*100000,3)) %>%
  mutate(travel_incidence_dot_log = log(travel_incidence_dot+0.02)) %>%
  mutate(Cases = replace_na(Cases, 0)) %>%
  mutate(Incidence = replace_na(Incidence, 0))
```


MAKE PREDICTIONS INTO INCIDENCE RATES AND SEPARATE 
```{r}
#Make predictions
df_pred_mod_final = 
  make_predictions(mod_final,df_yr) %>%
  mutate(pred_inc = pred/Pop*100000,
         pred_lwr_inc = pred_lwr/Pop*100000,
         pred_upr_inc = pred_upr/Pop*100000) %>%
  rename(`Local Incidence` = Incidence)
```


MANIPULATE FOR THE LEGEND
```{r}
t1 = df_pred_mod_final %>%
  select(Country, Year, pred_inc) %>%
  mutate(`Local Incidence` = "Predicted") %>%
  rename(value = pred_inc)

t2 = df_pred_mod_final %>%
  select(Country, Year, `Local Incidence`) %>%
  mutate(x = paste0(Country, " (Observed)")) %>%
  rename(value = `Local Incidence`) %>%
  rename(`Local Incidence` = x)

df_plot_test = rbind(t1,t2) %>%
           filter(Country %in% c("Cuba", "Haiti"))

df_plot_tr = rbind(t1,t2) %>%
         filter(Country %in% c("Puerto Rico", "Dominican Republic", "Jamaica"))

rm(t1,t2)


df_pred_mod_final_tr = df_pred_mod_final %>% 
         filter(Country %in% c("Puerto Rico", "Dominican Republic", "Jamaica"))

df_pred_mod_final_test = df_pred_mod_final %>% 
         filter(Country %in% c("Cuba", "Haiti"))


write.csv(df_pred_mod_final, "data_output/df_predictions_final.csv")
```


COMPARE PREDICTED INCIDENCE AND CASES WITH REPORTED CASES IN THE REGION AS A BOXPLOT
```{r}
#list of countries to remove that ever had a population under 1M
t = case_paho_yr %>%
  filter(Population < 1000000 | Country %in% c("United States of America", "Canada")) %>%
  distinct(Country)

inc_comp = case_paho_yr %>% 
  left_join(df_pred_mod_final, by = c("Country", "Year")) %>%
  select(Country, Year, Incidence, pred_inc, pred_lwr_inc, pred_upr_inc, 
         Cases.x, Cases.y, pred, pred_lwr, pred_upr)  %>%
  mutate(Type = if_else(is.na(pred_inc),
                        "Observed", "Predicted"),
         level = if_else(is.na(pred_inc), Incidence, pred_inc)) %>%
  anti_join(t) #remove countries with less than 1M Pop
  

group_ordered <- with(inc_comp,                       # Order boxes by median
                      reorder(Country,
                              level,
                              median))

inc_comp = inc_comp %>%
  mutate(Country = factor(Country, levels = levels(group_ordered))) %>%
  select(-Incidence) %>%
  rename(`Yearly Incidence (cases/pop*100k)` = level)

# inc_comp2 = inc_comp %>%
#   select(Country, Year, Incidence, pred_inc) %>%
#   rename(Observed = Incidence,
#          Predicted = pred_inc) %>%
#   pivot_longer(cols = c(-Country, -Year), names_to = "Type", values_to = "Incidence (2022)") %>%
#   mutate(Country = factor(Country, levels = level))


# #flip the x and y axis
 p_inc_comp = ggplot(inc_comp %>% filter(Year == 2022), aes(y = `Yearly Incidence (cases/pop*100k)`, x = Country, color = Type, fill = Type)) +
   geom_col(position = "dodge", alpha = .5) +
   scale_fill_manual(values = c("Observed" = obs_color,
                                "Predicted" = pred_color)) +
     scale_color_manual(values = c("Observed" = obs_color,
                                   "Predicted" = pred_color)) +
  theme_classic() +
  theme(axis.title.x = element_blank(),
         axis.text.x = element_text(angle = 45, vjust = 1),
         legend.title = element_blank(),
        legend.position = "bottom")

 p_inc_comp

p_inc_comp3 = ggplot(inc_comp) +
  geom_boxplot(aes(x = `Yearly Incidence (cases/pop*100k)`, y = Country,
                   color = Type, fill = Type),
               outlier.size = 3, alpha = .5) +
  geom_text(x = 1400, y = "Cuba", label = "2022", color = obs_color, size = label_text) +
  scale_fill_manual(values = c("Observed" = obs_color,
                               "Predicted" = pred_color)) +
  scale_color_manual(values = c("Observed" = obs_color,
                                  "Predicted" = pred_color)) +
  theme_classic() +
  ggtitle("C") +
  theme(axis.title.y = element_blank(),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 8),
        plot.title = element_text(face = "bold"),
        legend.title = element_blank(),
        legend.position = "bottom") +
  guides(label = NA,
         color = guide_legend(override.aes = aes(shape = 15, size = 4)))

p_inc_comp3

```

PLOT THE TIMESERIES
```{r}
#manipulate back to original with no country colors
df_plot_tr = df_plot_tr %>%
  mutate(`Local Incidence` = if_else(`Local Incidence` == "Predicted", 
                                     `Local Incidence`, 
                                     "Observed"))

df_plot_test = df_plot_test %>%
  mutate(`Local Incidence` = if_else(`Local Incidence` == "Predicted", 
                                     `Local Incidence`, 
                                     "Observed"))


#Plot Timeseries removing the local cases for all countries
ts_fun = function(df1, df2, ymax) {
  ggplot() +
    geom_line(data =df1, aes(x = Year, y = value, color = `Local Incidence`), size = 1) +
    geom_ribbon(data = df2, aes(x = Year, y = pred_inc, ymin = pred_lwr_inc, ymax = pred_upr_inc, group = 1), fill = "darkred",alpha = 0.2) +
    labs(color = "Local Incidence") +
    coord_cartesian(ylim = c(0,ymax)) +
    facet_wrap(~Country, ncol = 1, strip.position = "top") +
    theme_classic() +
    #scale_color_manual(values = c(country_colors3)) +
    scale_color_manual(values = fig_colors) +
    ylab("Incidence (cases/pop*100k)") +
    theme(axis.title.x = element_blank(),
          legend.position = "bottom",
          legend.title = element_blank(),
          axis.title = element_text(size = axis_title),
          axis.text = element_text(size = axis_text),
          strip.background = element_blank(),
          strip.text = element_text(hjust = 0),
          plot.title = element_text(size = title,
                                    face = face)
          ) #+
   # guides(color=guide_legend(nrow=2,byrow=TRUE))
}

p_final_tr_ts = ts_fun(df1 = df_plot_tr, df2 = df_pred_mod_final_tr, ymax = 1000) + ggtitle("A") + theme(legend.position = "none")

p_final_test_ts <- ts_fun(df1 = df_plot_test, df2 = df_pred_mod_final_test, ymax = 2000)  + ggtitle("B") 


p_final_tr_ts/p_final_test_ts + plot_layout(guides = "collect")


```

FORMAT FINAL PLOT
```{r}
#p_final_est/p_final_scat/p_inc_comp|p_final_tr_ts/p_final_test_ts


#p_final_est/p_final_scat|p_final_test_ts/p_inc_comp

p_final_fig = p_final_tr_ts/p_final_test_ts|p_inc_comp3

p_final_fig

ggsave(plot = p_final_fig, filename = "figures/p_final_fig.svg", width = 8, height = 10, units = "in")
ggsave(plot = p_final_fig, filename = "figures/p_final_fig.png", width = 8, height = 10, units = "in")
```
